Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-05-25 cvx-flats-neurips21/cvx-flats.tex
Minor changes; resolving comments and such

The following lemma yields consistency bounds when combined with Proposition~\ref{thm:consistent-implies-indir-elic}.\raft{Minor writing point: we capitalize lemma when it's a proper noun, hence ``this lemma'' vs ``Lemma 1''.  I know, weird.}
% Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop} (in Appendix~\ref{app:implications-of-bounds}), which 
A similar result is found in \citet[Theorem 9]{agarwal2015consistent}, which bounds the dimension of level sets of a single-valued $\prop{L}$.
Theorem~\ref{thm:convex-flats-inf-dim} instead bounds the dimension of flats contained in the level sets, an additional power which we leverage in our examples.

\begin{restatable}{theorem}{convexflats}  \label{thm:convex-flats-inf-dim}
  Let $\Gamma:\P \toto \reals^d$ be (directly) elicited by $L \in \Lcvx_d$ for some $d\in\mathbb{N}$.
  Let $\Y$ be either a finite set, or $\Y = \reals$, in which case we assume each $p\in\P$ admits a Lebesgue density supported on the same set for all $p\in\P$.%
  \footnote{This assumption is largely for technical convenience, to ensure that $\V_{u,p}$ does not depend on $p$.
    Any such assumption would suffice, and we suspect even that condition can be relaxed.}
  For all $u\in\range\Gamma$ and $p\in\Gamma_u$, there is some $V_{u,p}:\Y\to\reals^d$ such that $p \in \zeros{V_{u,p}} \subseteq \Gamma_u$.
\raft{Suggestion: let's ditch $V_{u,p}$ in the statement, and instead say ``there is some $d$-flat $F$ such that $p \in F \subseteq \Gamma_u$'', since (a) that's the way we write it everywhere else, (b) we justify the story that our main tool is about $d$-flats, and (c) it becomes clearer why the two Corollaries could follow.  Am I missing a reason to keep $V_{u,p}$?}
\end{restatable}
\begin{proof}[Finite case]
  We will prove the result for the case of finite $\Y$, and defer the $Y=\reals$ case to Appendix~\ref{app:implications-of-bounds}.
  As $L$ is convex and elicits $\Gamma$, we have $u \in \Gamma(p) \iff \vec 0 \in \partial \exploss{L}{u}{p}$. 
  With $\Y$ finite, this is additionally equivalent to $\vec 0 \in \oplus_y p_y \partial L(u,y)$, where $\oplus$ denotes the Minkowski sum~\citep[Theorem 4.1.1]{hiriart2012fundamentals}.\footnote{$\partial$ represents the subdifferential $\partial f(x) = \{z : f(x') - f(x) \geq \inprod{z}{x'-x}\; \forall x' \}$.}
  Expanding, we have $\oplus_y p_y \partial L(u,y) = \{ \sum_{y\in\Y} p_y x_y \mid x_y \in \partial L(u,y) \; \forall y\in\Y\}$, and thus $W p = \sum_y p_y x_y = \vec 0$ where $W = [x_1, \ldots, x_n] \in \reals^{d\times n}$; cf.~\cite[$\mathbf{A}^m$ in Theorem 16]{ramaswamy2016convex}.
  Let $V_{u,p} : \Y \to \reals^d, y \mapsto W_y$ be the function encoding the columns of $W$.
  Observe that $\E_p V_{u,p} = \vec 0$.				 
  We take the flat $F := \zeros{V_{u,p}}$, and have $p \in F$ by construction.
  To see $F \subseteq \Gamma_u$, from the chain of equivalences above, we have for any $q\in\P$ that $q \in \zeros{V_{u,p}} \implies \vec 0 \in \partial \E_q L(u,Y) \implies u \in \Gamma(q) \implies q \in \Gamma_u$.
\end{proof}

@@

\begin{scope}
\clip (h) -- (m) -- (l) -- (h);
\foreach \c in {0,0.1,...,1}
{
	\pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
	\pgfmathsetmacro\bb{0.5*\c}
	\coordinate (A) at ($(\aa,1-\aa,0)$);
	\coordinate (B) at ($(1-\aa,\aa,0)$);
	\coordinate (C) at ($(\bb,\bb,1-2*\bb)$);
	\draw[myblue,thin] (C) parabola (A);
	\draw[myblue,thin] (C) parabola (B);
}
\def\c{0.6}
\pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
\pgfmathsetmacro\bb{0.5*\c}
\coordinate (A) at ($(\aa,1-\aa,0)$);
\coordinate (B) at ($(1-\aa,\aa,0)$);
\coordinate (C) at ($(\bb,\bb,1-2*\bb)$);
\coordinate (Cleft) at ($(\bb + 2,\bb,1-2*\bb)$);
\coordinate (Cright) at ($(\bb,\bb + 2,1-2*\bb)$);
\coordinate (centerdot) at (0.21,0.445, 0.345);
\draw[myblue, very thick] (C) parabola (A);
\draw[myblue, very thick] (C) parabola (B);
\draw[orange, dashed,shorten >= -3cm, shorten <=-3cm] (1/10,9/10,0) -- (centerdot);
\node at (centerdot) {$\bullet$};
\end{scope}
\def\c{0.6}
\pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
\pgfmathsetmacro\bb{0.5*\c}
\node (levset) at (0.5,0.5,-.3) {$\Gamma_{0.6}$};
\draw[-latex,thin] (levset) edge[in=270,out=180] ($(\aa,1-\aa,0)$);
\draw[-latex,thin] (levset) edge[in=270,out=0] ($(1-\aa,\aa,0)$);
\placefiglabel{(Q4)}
\end{tikzpicture}
%\\
%\hline
\end{tabular}
\caption{Example properties for each of the quadrants. 
In each Quadrant, we take $\bullet$ to be the distribution $p = (0.3, 0.3, 0.4)$ according to the left, top, and right outcomes respectively. 
In Q1, we cannot fit a $1$-flat (line) through $p$ without leaving the level set $\Gamma_\bot$.  
Therefore, there is no $1$-dimensional consistent convex surrogate for abstain by Theorem~\ref{thm:convex-flats-inf-dim}.  
(Q2) is similar to (Q1) but has no target loss since it is not elicitable.
In (Q3), we can fit a $1$-flat through any $p \in \simplex$ for least squares, eliciting expected value.
In (Q4), we look at the variance, and take the distribution $p = \bullet \in \Gamma_{0.6}$.
Since $\Gamma_{0.6}$ (solid orange) is curved, we cannot draw a $1$-flat contained in the level set, and conclude there is not a $1$-dimensional consistent convex surrogate w.r.t. the variance.
}\label{fig:quadrants-simplices}
\end{SCfigure}

@@

\newpage
\appendix
\section{Consistency implies indirect elicitation}\label{sec:consis-implies-indir}

>>>>

The following lemma yields consistency bounds when combined with Proposition~\ref{thm:consistent-implies-indir-elic}. %\raft{Minor writing point: we capitalize lemma when it's a proper noun, hence ``this lemma'' vs ``Lemma 1''.  I know, weird.}
% Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop} (in Appendix~\ref{app:implications-of-bounds}), which 
A similar result is found in \citet[Theorem 9]{agarwal2015consistent}, which bounds the dimension of level sets of a single-valued $\prop{L}$.
Theorem~\ref{thm:convex-flats-inf-dim} instead bounds the dimension of flats contained in the level sets, an additional power which we leverage in our examples.

\begin{restatable}{theorem}{convexflats}  \label{thm:convex-flats-inf-dim}
  Let $\Gamma:\P \toto \reals^d$ be (directly) elicited by $L \in \Lcvx_d$ for some $d\in\mathbb{N}$.
  Let $\Y$ be either a finite set, or $\Y = \reals$, in which case we assume each $p\in\P$ admits a Lebesgue density supported on the same set for all $p\in\P$.%
  \footnote{This assumption is largely for technical convenience, to ensure that $\V_{u,p}$ does not depend on $p$.
    Any such assumption would suffice, and we suspect even that condition can be relaxed.}
  For all $u\in\range\Gamma$ and $p\in\Gamma_u$, there is some $d$-flat $F$ such that $p \in F \subseteq \Gamma_u$.
 % For all $u\in\range\Gamma$ and $p\in\Gamma_u$, there is some $V_{u,p}:\Y\to\reals^d$ such that $p \in \zeros{V_{u,p}} \subseteq \Gamma_u$.
\raft{Suggestion: let's ditch $V_{u,p}$ in the statement, and instead say ``there is some $d$-flat $F$ such that $p \in F \subseteq \Gamma_u$'', since (a) that's the way we write it everywhere else, (b) we justify the story that our main tool is about $d$-flats, and (c) it becomes clearer why the two Corollaries could follow.  Am I missing a reason to keep $V_{u,p}$? \jessie{I like this change; kept the old version in a comment, but no reason I can think of to keep $V_{u,p}$}}
\end{restatable}
\begin{proof}[Finite case]
  We will prove the result for the case of finite $\Y$, and defer the $\Y=\reals$ case to Appendix~\ref{app:implications-of-bounds}.
  As $L$ is convex and elicits $\Gamma$, we have $u \in \Gamma(p) \iff \vec 0 \in \partial \exploss{L}{u}{p}$. 
  With $\Y$ finite, this is additionally equivalent to $\vec 0 \in \oplus_y p_y \partial L(u,y)$, where $\oplus$ denotes the Minkowski sum~\citep[Theorem 4.1.1]{hiriart2012fundamentals}.\footnote{$\partial$ represents the subdifferential $\partial f(x) = \{z : f(x') - f(x) \geq \inprod{z}{x'-x}\; \forall x' \}$.}
  Expanding, we have $\oplus_y p_y \partial L(u,y) = \{ \sum_{y\in\Y} p_y x_y \mid x_y \in \partial L(u,y) \; \forall y\in\Y\}$, and thus $W p = \sum_y p_y x_y = \vec 0$ where $W = [x_1, \ldots, x_n] \in \reals^{d\times n}$; cf.~\cite[$\mathbf{A}^m$ in Theorem 16]{ramaswamy2016convex}.
  Let $V_{u,p} : \Y \to \reals^d, y \mapsto W_y$ be the function encoding the columns of $W$.
  Observe that $\E_p V_{u,p} = \vec 0$.				 
  We take the flat $F := \zeros{V_{u,p}}$, and have $p \in F$ by construction.
  To see $F \subseteq \Gamma_u$, from the chain of equivalences above, we have for any $q\in\P$ that $q \in \zeros{V_{u,p}} \implies \vec 0 \in \partial \E_q L(u,Y) \implies u \in \Gamma(q) \implies q \in \Gamma_u$.
\end{proof}

@@

\begin{scope}
\clip (h) -- (m) -- (l) -- (h);
\foreach \c in {0,0.1,...,1}
{
	\pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
	\pgfmathsetmacro\bb{0.5*\c}
	\coordinate (A) at ($(\aa,1-\aa,0)$);
	\coordinate (B) at ($(1-\aa,\aa,0)$);
	\coordinate (C) at ($(\bb,\bb,1-2*\bb)$);
	\draw[myblue,thin] (C) parabola (A);
	\draw[myblue,thin] (C) parabola (B);
}
\def\c{0.6}
\pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
\pgfmathsetmacro\bb{0.5*\c}
\coordinate (A) at ($(\aa,1-\aa,0)$);
\coordinate (B) at ($(1-\aa,\aa,0)$);
\coordinate (C) at ($(\bb,\bb,1-2*\bb)$);
\coordinate (Cleft) at ($(\bb + 2,\bb,1-2*\bb)$);
\coordinate (Cright) at ($(\bb,\bb + 2,1-2*\bb)$);
\coordinate (centerdot) at (0.21,0.445, 0.345);
\draw[myblue, very thick] (C) parabola (A);
\draw[myblue, very thick] (C) parabola (B);
\draw[orange, dashed,shorten >= -3cm, shorten <=-3cm] (1/10,9/10,0) -- (centerdot);
\node at (centerdot) {$\bullet$};
\end{scope}
\def\c{0.6}
\pgfmathsetmacro\aa{0.5*(sqrt(1-\c)+1)}
\pgfmathsetmacro\bb{0.5*\c}
\node (levset) at (0.5,0.5,-.3) {$\Gamma_{0.6}$};
\draw[-latex,thin] (levset) edge[in=270,out=180] ($(\aa,1-\aa,0)$);
\draw[-latex,thin] (levset) edge[in=270,out=0] ($(1-\aa,\aa,0)$);
\placefiglabel{(Q4)}
\end{tikzpicture}
%\\
%\hline
\end{tabular}
\caption{Example properties for each of the quadrants. 
In each Quadrant, we take $\bullet$ to be the distribution $p = (0.3, 0.3, 0.4)$ according to the left, top, and right outcomes respectively. 
In (Q1) and (Q2), we cannot fit a $1$-flat (line) through $p$ without leaving the level set $\Gamma_\bot$.  
Therefore, there is no $1$-dimensional consistent convex surrogate for abstain by Theorem~\ref{thm:convex-flats-inf-dim}.  
In (Q3), we can fit a $1$-flat through any $p \in \simplex$ for least squares, eliciting expected value.
In (Q4), the level sets of the variance are curved, and therefore we cannot draw a $1$-flat contained in the level set, and conclude there is not a $1$-dimensional consistent convex surrogate w.r.t. the variance.
}\label{fig:quadrants-simplices}
\end{SCfigure}

@@

\newpage
\section*{Checklist}

\begin{enumerate}

	\item For all authors...
	\begin{enumerate}
		\item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
		\answerYes{}
		\item Did you describe the limitations of your work?
		\answerYes{}
		\item Did you discuss any potential negative societal impacts of your work?
		\answerYes{} \textbf{See discussion in the Introduction; it is hard to forecast impacts of theoretical work, but we try to explain some impacts.}
		\item Have you read the ethics review guidelines and ensured that your paper conforms to them?
		\answerYes{}
	\end{enumerate}

	\item If you are including theoretical results...
	\begin{enumerate}
		\item Did you state the full set of assumptions of all theoretical results?
		\answerYes{}\textbf{Sometimes details are deferred to Appendix, but we make an effort to explicitly state when that happens.}
		\item Did you include complete proofs of all theoretical results?
		\answerYes{}\textbf{Sometimes proofs are deferred to Appendix, but we make an effort to explicitly state when that happens.}
	\end{enumerate}

	\item If you ran experiments...
	\begin{enumerate}
		\item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
		\answerNA{}
		\item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
		\answerNA{}
		\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
		\answerNA{}
		\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
		\answerNA{}
	\end{enumerate}

	\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
	\begin{enumerate}
		\item If your work uses existing assets, did you cite the creators?
		\answerNA{}
		\item Did you mention the license of the assets?
		\answerNA{}
		\item Did you include any new assets either in the supplemental material or as a URL?
		\answerNA{}
		\item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
		\answerNA{}
		\item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
		\answerNA{}
	\end{enumerate}

	\item If you used crowdsourcing or conducted research with human subjects...
	\begin{enumerate}
		\item Did you include the full text of instructions given to participants and screenshots, if applicable?
		\answerNA{}
		\item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
		\answerNA{}
		\item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
		\answerNA{}
	\end{enumerate}

\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix
\section{Consistency implies indirect elicitation}\label{sec:consis-implies-indir}
